\Large \bf Vigor: Toward Carrier-Grade Software Network Functions

### 2016-10-27 doc/paper/paper.tex
Missing reference; wrap code listings into figures with captions.

\lstset{language=C,
    basicstyle=\footnotesize\ttfamily,
    keywordstyle=\color{myblue}\ttfamily,
    stringstyle=\color{mygreen}\ttfamily,
    commentstyle=\color{gray}\ttfamily,
    keepspaces=true,
    tabsize=2,
    otherkeywords={bool}
}

@@

\subsection*{Abstract}
The objective of this work is to enable developers to productively write software-based data plane network functions that simultaneously meet the highest standards of reliability, security, and performance. We present a verification approach for formally proving properties of stateful network functions, in a way that is practical both in terms of developer effort and the performance of the resulting data plane software. Unlike alternative approaches that require drastic changes in the way network functions are developed, our approach allows the use of C/C++ together with popular frameworks like DPDK. To the best of our knowledge, our work is the first to both be practical and formally verify stateful network functions. We introduce a new ``speculative'' form of symbolic execution, which achieves scalability to real-world network functions by leveraging insights related to the structure and cyclomatic complexity of data plane software. This technique dovetails with a new library of verified components that are commonly used in network functions. We illustrate our approach by developing a middlebox that is verified to never crash, to not leak memory, be free of several classes of security vulnerabilities, and whose performance is on par with other NATs. %\footnote{Disclaimer: we use acronym NAT in this paper to refer to a more narrow kind - Dynamic Source NAPT (network address-port translation~\cite{rfc3022}).}. 
We view this as an important milestone on the road to complete verification of all software-based network functions.

@@

We achieve both scalable verification and minimal developer effort by leveraging two insights into the verification of data plane software:
\begin{itemize}
    \item {\bf Canonical data structures}: Unlike general-purpose software, data plane software employs a small number of commonly used network data structures (hash table, least-prefix-match, rings, etc.) and components. These lend themselves naturally to being abstracted into a library. They can be formally verified a priori, and their proofs then automatically leveraged for proving properties of the code that uses them.
    \item {\bf Hybrid verification}: While typically one picks a verification technique and applies it throughout, any {\em one} technique works well for {\em some} of the code and not for other. However, data plane software offers us the opportunity to advantageously employ symbolic execution for high-level code (which is mostly stateless and has short execution paths) and verification based on separation logic to the library of components/data structures (which is highly stateful).

@@

\begin{lstlisting}
#define CAP 512

@@

  while(1) {
    if (end != (begin - 1) || 
        !(end == CAP - 1 && begin == 0)) {
      if (receive_packet(&packets[end]) && 
          packets[end].port != 9)
        end = (end + 1)%CAP;
    }
    if (end != begin && can_send_packet()) {
      send_packet(&packets[begin]);
      begin = (begin + 1)%CAP;
    }
  }
  return 0;
}

\end{lstlisting}

@@

{\bf Step 1: Add loop invariants.} Infinite (or very long) loops can be simplified using loop invariants~\cite[\S~2.1]{cormen2009introduction}. These are annotations that inform the SEE about invariant properties of the loop, such as ``no element of \code{packets} between \code{begin} and \code{end} has destination port equal to 9'', and are a good match for NF-style loops.  We added to KLEE support for havoc-ing~\cite{boogie} automatically, to take advantage of such invariants and essentially reduce the infinite loop to a single abstract iteration that can be reasoned about in one pass.

@@

\begin{lstlisting}
int main(int argc, char** argv) {
  struct packet p;
  struct ring *r = ring_create(CAP);
  if (!r) return 1;
  while(LOOP(1))
  {
    loop_iteration_begin(&r);
    if (!ring_full(r))
      if (receive_packet(&p) && p.port != 9)
        ring_push_back(r, &p);
    if (!ring_empty(r) && can_send_packet()) {
      ring_pop_front(r, &p);
      VIGOR_CHECK(p.port != 9);
      send_packet(&p);
    }
    loop_iteration_end(&r);
  }
  return 0;
}
\end{lstlisting}

\begin{lstlisting}
#define CAP 512
#define RING_CAPACITY 512

@@

static bool packet_constraints(struct packet* p) {
  return p->port != 9;
}
\end{lstlisting}

@@

Suppose we had an overapproximate model:
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
}
\end{lstlisting}
Then the \code{ASSERT} from the previous listing would fail.

@@

\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p);
/*@ requires ringp(r, ?lst,
                   ?property, ?capacity) &*&
             lst != nil &*&
             packetp(p, _); @*/
/*@ ensures ringp(r, tail(lst),
                  property, capacity) &*&
            packetp(p, head(lst)) &*&
            true == property(head(lst)); @*/
\end{lstlisting}

@@

\begin{lstlisting}
struct ring* arg1;
struct packet arg2;

@@

/*@ assert((false == (9 == arg2.port))); @*/
\end{lstlisting}

Suppose we had an underapproximate model:
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  p->port = 0;
}
\end{lstlisting}
then the last line in the call sequence check will look like:
\begin{lstlisting}
/*@ assert(0 == arg2.port); @*/
\end{lstlisting}
And the formal verification will fail, because specification admits a wider
range of possible outputs.

>>>>

\lstset{language=C,
    basicstyle=\footnotesize\ttfamily,
    keywordstyle=\color{myblue}\ttfamily,
    stringstyle=\color{mygreen}\ttfamily,
    commentstyle=\color{gray}\ttfamily,
    keepspaces=true,
    tabsize=2,
    otherkeywords={bool},
    numbers=left,
    numbersep=1pt,
    numberstyle=\tiny\color{gray}
}

@@

\subsection*{Abstract}
The objective of this work is to enable developers to productively write
software-based data plane network functions that simultaneously meet the highest
standards of reliability, security, and performance. We present a verification
approach for formally proving properties of stateful network functions, in a way
that is practical both in terms of developer effort and the performance of the
resulting data plane software. Unlike alternative approaches that require
drastic changes in the way network functions are developed, our approach allows
the use of C/C++ together with popular frameworks like DPDK. To the best of our
knowledge, our work is the first to both be practical and formally verify
stateful network functions. We introduce a new ``speculative'' form of symbolic
execution, which achieves scalability to real-world network functions by
leveraging insights related to the structure and cyclomatic complexity of data
plane software. This technique dovetails with a new library of verified
components that are commonly used in network functions. We illustrate our
approach by developing a NAT box that is verified to never crash, to not leak
memory, be free of several classes of security vulnerabilities, and whose
performance is on par with other NATs. %\footnote{Disclaimer: we use acronym NAT in this paper to refer to a more narrow kind - Dynamic Source NAPT (network address-port translation~\cite{rfc3022}).}. 
We view this as an important milestone on the road to complete verification of all software-based network functions.

@@

We achieve both scalable verification and minimal developer effort by leveraging two insights into the verification of data plane software:
\begin{itemize}
    \item {\bf Canonical data structures}: Unlike general-purpose software, data plane software employs a small number of commonly used network data structures (hash table, least-prefix-match, rings, etc.) and components. These lend themselves naturally to being abstracted into a library. They can be formally verified a priori, and their proofs then automatically leveraged for proving properties of the code that uses them.
    \item {\bf Hybrid verification}: While typically one picks a verification
      technique and applies it throughout, no {\em single} technique works well for {\em all} of the code. However, data plane software offers us the opportunity to advantageously employ symbolic execution for high-level code (which is mostly stateless and has short execution paths) and verification based on separation logic to the library of components/data structures (which is highly stateful).

@@

\begin{figure}[h!]
\begin{lstlisting}
#define CAP 512

@@

  while(1) {
    if (end != (begin - 1) || 
        !(end == CAP - 1 && begin == 0)) {
      if (receive_packet(&packets[end]) && 
          packets[end].port != 9)
        end = (end + 1)%CAP;
    }
    if (end != begin && can_send_packet()) {
      send_packet(&packets[begin]);
      begin = (begin + 1)%CAP;
    }
  }
  return 0;
}
\end{lstlisting}
  \caption{Original implementation of the Drop protocol}
  \label{lst:orig}
\end{figure}

@@

{\bf Step 1: Add loop invariants.} Infinite (or very long) loops can be simplified using loop invariants~\cite[\S~2.1]{cormen2009introduction}. These are annotations that inform the SEE about invariant properties of the loop, such as ``no element of \code{packets} between \code{begin} and \code{end} has destination port equal to 9'', and are a good match for NF-style loops.  We added to KLEE support for havoc-ing~\cite{barnett2005boogie} automatically, to take advantage of such invariants and essentially reduce the infinite loop to a single abstract iteration that can be reasoned about in one pass.

@@

\begin{figure}[h!]
\begin{lstlisting}
int main(int argc, char** argv) {
  struct packet p;
  struct ring *r = ring_create(CAP);
  if (!r) return 1;
  while(LOOP(1))
  {
    loop_iteration_begin(&r);
    if (!ring_full(r))
      if (receive_packet(&p) && p.port != 9)
        ring_push_back(r, &p);
    if (!ring_empty(r) && can_send_packet()) {
      ring_pop_front(r, &p);
      VIGOR_CHECK(p.port != 9);
      send_packet(&p);
    }
    loop_iteration_end(&r);
  }
  return 0;
}
\end{lstlisting}
  \caption{Vigor implementation of Drop protocol.}
  \label{lst:vigor}
\end{figure}

For the symbolic execution and future validation, we need also to customize the
general-purpose packet ring with the following definitions:

\begin{figure}
\begin{lstlisting}
#define CAP 512
#define RING_CAPACITY CAP

@@

static bool packet_constraints(struct packet* p) {
  return p->port != 9;
}
\end{lstlisting}
  \caption{Custom definitions capturing the desired property.}
  \label{lst:customizations}
\end{figure}

@@

Suppose we had an overapproximate model:
\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  FILL_SYMBOLIC(p, sizeof(struct packet),
                "popped_packet");
}
\end{lstlisting}
  \caption{An extract from an overapproximate ring model.}
  \label{lst:overapproximate}
\end{figure}
Then the \code{ASSERT} from the previous listing would fail.

@@

\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p);
/*@ requires ringp(r, ?lst,
                   ?property, ?capacity) &*&
             lst != nil &*&
             packetp(p, _); @*/
/*@ ensures ringp(r, tail(lst),
                  property, capacity) &*&
            packetp(p, head(lst)) &*&
            true == property(head(lst)); @*/
\end{lstlisting}
  \caption{A formal contract for the ring\_pop\_front function.}
  \label{lst:contract}
\end{figure}

@@

\begin{figure}[h!]
\begin{lstlisting}
struct ring* arg1;
struct packet arg2;

@@

/*@ assert((false == (9 == arg2.port))); @*/
\end{lstlisting}
  \caption{A call sequence, produced by SEE.}
  \label{lst:call-sequence}
\end{figure}

Suppose we had an underapproximate model:
\begin{figure}[h!]
\begin{lstlisting}
void ring_pop_front(struct ring* r,
                    struct packet* p) {
  p->port = 0;
}
\end{lstlisting}
  \caption{An underapproximate model fragment.}
  \label{lst:underapproximate}
\end{figure}
then the last line in the call sequence check will look like:
\begin{lstlisting}
/*@ assert(0 == arg2.port); @*/
\end{lstlisting}
And the formal verification will fail, because specification admits a wider
range of possible outputs.
